Bugs and Fixes
========================================================

## Add 1: nls.lm as the standard optimization procedure. 


## Bug 1: Anastasiia, 27/03/2013

reason: k2=0 ! How to solve it??

```{r}

 WW <- mkinmod.full(
         c1 = list(
       time = c(    0,     0,     3,     3,     7,     7,    14,    14,    21,    21,    30,    30,    45,    45,    60,    60,    90,    90,   120,   120),
    residue = c(75.04, 72.88, 74.00, 72.69, 62.49, 69.02, 62.05, 63.74, 58.88, 56.20, 52.70, 50.39, 40.20, 40.73, 29.43, 29.89, 29.16, 30.70, 25.41, 26.38),
     weight = c(    1,     1,     1,     1,     1,     1,     1,     1,     1,     1,     1,     1,     1,     1,     1,     1,     1,     1,     1,     1),
                   sink  = TRUE,
       type = "DFOP",
         k1 = list(ini   = 0.1000,
                   fixed = 0,
                   lower = 0.0,
                   upper = Inf),
         k2 = list(ini   = 0.0100,
                   fixed = 0,
                   lower = 0.0,
                   upper = Inf),
          g = list(ini   = 0.3000,
                   fixed = 0,
                   lower = 0.0,
                   upper = 1),
         M0 = list(ini   = 75.04,
                   fixed = 0,
                   lower = 0.0,
                   upper = Inf)))


#
# Fit and optimizer
#

  Fit    <- IRLSkinfit.full(
            WW,
               plot      = TRUE,
               quiet     = TRUE,
               ctr       = kingui.control(
                               method = 'solnp',
                            submethod = 'Port',
                              maxIter = 100,
                            tolerance = 1E-06,
                            odesolver = 'lsoda'),
            irls.control = list(
                              maxIter = 10,
                            tolerance = 0.001))

```




In case the library cannot be loaded.
```{r env,eval=FALSE}
.libPaths( c( .libPaths(), "M:/Personal Data/R/win-library/2.15/") ) ## why does it not work????????????????
library("KineticEval",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("deSolve",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("rootSolve",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("minpack.lm",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("coda",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("FME",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("mkin",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("numDeriv",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("optimx",lib.loc="M:/Personal Data/R/win-library/2.15/")

## Not required
library("quadprog",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("BB",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("ucminf",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("Rcgmin",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("Rvmmin",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("Rcpp",lib.loc="M:/Personal Data/R/win-library/2.15/")
library("minqa",lib.loc="M:/Personal Data/R/win-library/2.15/")

load("E:/KinEvalGit/exchange/andrew.rda")
Sys.getenv()
```

Now assuming the sourcefunctions/packages are loaded.


## Bug 1: Cannot reach the optimal solution

```{r bug1}
## 
mkinmodini <- mkinmod.full(Parent=list(type="SFO",to="Metab"),
                           Metab=list(type="SFO",M0=list(ini=0,fixed=0,lower=0,upper=Inf)),data=andrew)
Fit <- mkinfit.full(mkinmodini,ctr= kingui.control(method = 'L-BFGS-B',submethod = 'Port'))
Fit.LM <- mkinfit.full(mkinmodini,ctr= kingui.control(method = 'Marq',submethod = 'Port'))
compare_multi_kinmod(mkinmodini,rbind(Fit$par,Fit.LM$par))
 
compare_multi_kinmod(mkinmodini,rbind(Fit$par,c(2.0633,0.2105,0.3033,0.8945,1)))
newmod1 <- update_kinmod(mkinmodini,newparms=Fit$par)$new_mkinmod
Fit1 <- mkinfit.full(newmod1,ctr= kingui.control(method = 'L-BFGS-B',submethod = 'Port'))
newmod2 <- update_kinmod(mkinmodini,newparms=c(2.0633,0.2105,0.3033,0.8945,1))$new_mkinmod
Fit2 <- mkinfit.full(newmod2,ctr= kingui.control(method = 'L-BFGS-B',submethod = 'Port'))
compare_multi_kinmod(mkinmodini,rbind(Fit$par,c(2.0633,0.2105,0.3033,0.8945,1),Fit2$par))
```

## Bug 2 Solved: the optimized parameters are not the same as specified in the model
```{r bug2}
## a possible bug that the optimized parameters are not the same as specified in the model
mkinmodini <- mkinmod.full(Parent=list(type="SFO",to="Metab"),
                           Metab=list(type="SFO"),data=andrew)
mkinmodiniFree <- mkinmod.full(Parent=list(type="SFO",to="Metab"),
                           Metab=list(type="SFO",M0=list(ini=0,fixed=0,lower=0,upper=Inf)),data=andrew)
Fit <- mkinfit.full(mkinmodini,ctr= kingui.control(method = 'Marq',submethod = 'Port'))
newmod1 <- update_kinmod(mkinmodini,newparms=Fit$par)$new_mkinmod
Fit1 <- mkinfit.full(newmod1,ctr= kingui.control(method = 'L-BFGS-B',submethod = 'Port'))
newmod2 <- update_kinmod(mkinmodini,newparms=c(2.0633,0.2105,0.3033,0.8945,1))$new_mkinmod
Fit2 <- mkinfit.full(newmod2,ctr= kingui.control(method = 'L-BFGS-B',submethod = 'Port'))
compare_multi_kinmod(mkinmodini,rbind(Fit$par,c(2.0633,0.2105,0.3033,0.8945,1),Fit2$par))

FitFree <-  mkinfit.full(mkinmodiniFree,ctr= kingui.control(method = 'Marq',submethod = 'Port'))

```
You can also embed plots, for example:

```{r fig.width=7, fig.height=6}

```

## Bug 3: warning of 0 in parameter values for Levenberg-Marquardt algorithm

```{r eval=FALSE}

mkinmodini <- mkinmod.full(Parent=list(type="SFO",to="Metab"),
                           Metab=list(type="SFO",M0=list(ini=0,fixed=0,lower=0,upper=Inf)),data=andrew)
Fit <- mkinfit.full(mkinmodini,ctr= kingui.control(method = 'L-BFGS-B',submethod = 'Port'))
Fit.LM <- mkinfit.full(mkinmodini,ctr= kingui.control(method = 'Marq',submethod = 'Port'))
compare_multi_kinmod(mkinmodini,rbind(Fit$par,Fit.LM$par))

```

It is not a bug, there is indeed 0 values in the parameters.

##  Bug 4:

When a compound has no metabolite and SINK is turned off, the program should automatically fix k=0.

## Improvements

The code can be similified by 

```{r}
a = 1
params = c()
params$a = 2
myfun <- function(params) {
  attach(params)
  print(a)
  detach(params)
}
myfun(params)

params$b <- list(c=3,d=4)
myfun2 <- function(params){
  # Assume params is a list of parameters from 
  # "params <- list(name1=value1, name2=value2, etc.)"
  for (v in 1:length(params)) assign(names(params)[v], params[[v]])
  # Use elements of params as parameters in a simulation
  print(a)
  print(b)
}
```
* http://stackoverflow.com/questions/871/why-is-git-better-than-subversion
* http://stackoverflow.com/questions/15187017/reduce-computing-time-for-reshape